Complete Algebraic File Synchronization

### 2015-12-24 2015/comparisons.tex
Improved text

\section{Comparisons}

\subsection{Liberal and conservative Reconciliation}

We consider the reconciliation algorithm described here an improvement over
the one derved during our previous research (see \cite{NREC})
as the previous algorithm not only fails to propagate all possible commands
wherever it is possible (that is, where it does not break the filesystem),
but the current algorithm is also simpler.
This is because the previous reconciliation algorithm excludes
commands from being propagated which must be preceded by a command that conflicts.

The former observation is supported by Berzan and Ramsey, who in \cite{CBNR} 
describe between different reconciliation policies.
The liberal (maximal) policy propagates all updates to all replicas where
the command does not break the filesystem, while a conservative policy
refrains from updating any node that is below a node with conflicting commands.
They show that the reconciliation algorithm in \cite{NREC} implements
an intermediate policy as one can easily construct two- or three-replica scenarios
where an update could clearly be propagated, but it is excluded.
To rephrase the example in \cite{CBNR}, consider the following two
update sequences that have been applied to replicas $\FS_A$ and $\FS_B$:
\begin{align*}
A&=\cbdaa{\nn}{\valdi}\cc\cbfaa{n}{\valf} \\
B&=\cbdaa{\nn}{\valdii}
\end{align*}
where $\parent(n)=\nn$ (\cite{NREC} did not require all directories to have the same value).
Clearly $\cbfaa{n}{\valf}$ could be applied to $\FS_B$, but it is not as
it must be preceded by creating the directory, which conflicts with the same operation in $B$
due to the different value.
(Introducing a third replica which has not changed at all further complicates the picture.)
The current algorithm no longer needs to specify that there can be no conflicts
on preceding commands, and, to use the above terminology, implements a fully liberal policy.

@@

As a test of the proposed reconciliation algorithm, 
we compare our definition of conflicting
updates to how conflicts are defined by Balasubramaniam and Pierce
in their state-based approach (\cite{BP}).
The update detector described in \cite{BP} provides a safe estimate of nodes
(paths) at which updates occurred.
It marks some nodes as \emph{dirty} in a way that we know that at non-dirty nodes
the filesystem have not changed between their original and current states
(see Definition 3.1.1 in \cite{BP}).
The \emph{dirty} marks are also up-closed, that is, all ancestor nodes of a dirty node
are also dirty (Fact 3.1.3 in \cite{BP}).
And finally, a conflict is detected between replicas $\FS_A$ and $\FS_B$ at node $n$
if $n$ is marked as dirty in both $\FS_A$ and $\FS_B$, and
$\FS_A(n)\neq\FS_B(n)$, and $n$ does not point to a directory in both replicas
($\FS_A(n)\not\in\setd \vee \FS_B(n)\not\in\setd$; see section 4.1 in \cite{BP}).

>>>>

\section{Related Work}

\subsection{Liberal and Conservative Reconciliation}

We consider the reconciliation algorithm described here an improvement over
the one derved during our previous research (\cite{NREC})
as the previous algorithm not only fails to propagate all possible commands
wherever it is possible (that is, where it does not break the filesystem),
but the current algorithm is also simpler.
This is because the previous reconciliation algorithm excludes
commands from being propagated which must be preceded by a command that conflicts.

The former observation is supported by Berzan and Ramsey, who in \cite{CBNR} 
describe different general reconciliation policies.
The liberal (maximal) policy propagates all updates to all replicas where
the update does not break the filesystem, while a conservative policy
refrains from updating any node that is below a node with conflicting commands.
They show that the reconciliation algorithm in \cite{NREC} implements
an intermediate policy as one can easily construct two- or three-replica scenarios
where an update could clearly be propagated, but it is excluded.
To rephrase the example in \cite{CBNR}, consider the following two
update sequences that have been applied to replicas $\FS_A$ and $\FS_B$:
\begin{align*}
A&=\cbdaa{\nn}{\valdi}\cc\cbfaa{n}{\valf} \\
B&=\cbdaa{\nn}{\valdii}
\end{align*}
where $\parent(n)=\nn$ (\cite{NREC} did not require all directories to have the same value).
Clearly $\cbfaa{n}{\valf}$ could be applied to $\FS_B$, but it is not as
it must be preceded by creating the directory, which conflicts with the same operation in $B$
due to the different value.
(Introducing a third replica which has not changed at all further complicates the picture.)
The current algorithm no longer needs to specify that there can be no conflicts
on preceding commands, and, to use the above terminology, implements a fully liberal policy.

@@

As a test of the proposed reconciliation algorithm, 
we compare our definition of conflicting
updates to how conflicts are defined by Balasubramaniam and Pierce
in their state-based approach (\cite{BP}).
The update detector they describe provides a safe estimate of nodes
(paths) at which updates occurred.
It marks some nodes as \emph{dirty} in a way that we know that at non-dirty nodes
the filesystems (replicas) have not changed between their common original state and current state
(see Definition 3.1.1 in \cite{BP}).
The \emph{dirty} marks are also up-closed, that is, all ancestor nodes of a dirty node
are also dirty (Fact 3.1.3 in \cite{BP}).
And finally, a conflict is detected between replicas $\FS_A$ and $\FS_B$ at node $n$
if $n$ is marked as dirty in both $\FS_A$ and $\FS_B$, and
$\FS_A(n)\neq\FS_B(n)$, and $n$ does not point to a directory in both replicas
($\FS_A(n)\not\in\setd \vee \FS_B(n)\not\in\setd$; see section 4.1 in \cite{BP}).

### 2015-12-24 2015/paper.tex
Improved text

% TODO
Conclusions go here -- same approach to other systems
\medskip

@@

Apart from constructing an algebra,
there are many other ways in which further work can extend the current results.
We think one of the most interesting extensions would be to
consider reconciling not only two, but more replicas in a single step and
prove the correctness of the algorithm proposed.
A related problem is to model cases where reconciliation cannot
complete fully (due to a system or network failure, for example)
or if only a subset of the replicas are reconciled,
both of which would result in a state where different replicas
have different common ancestors, that is,
the updates specific to the replicas start from different points
in the update history of the filesystem.
Identifying the information one needs to keep to perform
update detection is such cases and devising a suitable algorithm
will be an essential part of any implementation of a reconciler
based on the ideas presented here.

An extension of these results could also entail relaxing
the simplifying assumptions we used.
Identifying which results depend on assuming that $|\setd|=1$
and devising updated proofs or algorithms,
or a pre-processing step that reconciles meta-information in directories
can enable the algebraic system or an implementation
to identify and resolve a wider range of conflicts.
Similarly, designing pre- and post-processors that
present certain changes not as removal and creation of a node,
but as a rename,
could contribute to the user-friendliness of the reconciler
and help avoid human error.

>>>>

% TODO
Conclusions go here --- same approach to other systems
\medskip

@@

Apart from constructing an algebra,
there are many other ways in which further work can extend the current results.
We think one of the most interesting extensions would be to
consider reconciling not only two, but more replicas in a single step and
prove the correctness and completeness of the algorithm proposed.
A related problem is to model cases where reconciliation cannot
complete fully (due to a system or network failure, for example)
or if only a subset of the replicas are reconciled,
both of which would result in a state where different replicas
have different common ancestors, that is,
the updates specific to the replicas start from different points
in the update history of the filesystem.
Identifying the information one needs to keep to perform
update detection in such cases and devising a suitable algorithm
would be beneficial for any implementation of a reconciler
based on the ideas presented here.

An extension of these results could also entail relaxing
the simplifying assumptions we used.
Identifying which results depend on the assumption that $|\setd|=1$
and devising updated proofs or algorithms,
or a pre-processing step that reconciles meta-information in directories
can enable the algebraic system or an implementation
to identify and resolve a wider range of conflicts.
Similarly, designing pre- and post-processors that
present certain changes not as removal and creation of a node,
but as a rename,
could contribute to the user-friendliness of the reconciler
and help avoid human error.

### 2015-12-24 2015/rec-complete.tex
Minor changes
More improvements and references
Further research, conflict comparison, renaming unrelated to incomparable
Improved text

\subsection{Completeness of Reconciliation}

% Reconciliation is maximal
% -------------------------
% This is where we're using |D|=1
% TODO C.f. strict / liberal etc. reconciliation in references

@@

\begin{myth}\label{rec_is_complete}
If $A$ and $B$ are simple sequences,
and we select a sequence $S\subset \amb$ so that
it contains a command $\cxynv$ for which $\cxynv\nindep\bma$,
then $B\cc S$ either breaks all filesystems $A$ and $B$ do not break
(that is, $S$ breaks all possible $\FS_B$ replicas),
or it overrides incompatible changes introduced in the replica by $B$.
\end{myth}
Overriding changes is not desirable as in these cases the reconciler
does not have enough information to decide whether to promote
the change introduced by $A$ over that introduced by $B$ or vice versa.
These cases should be treated as conflicts and
the affected nodes or commands marked for review
by a human or a different system.
\begin{proof}
The proof is similar to the ones we have seen above.
Without loss of generality, we assume that $\cxynv$ is the first command in $S$
that is not independent of $\bma$.
If so, we can split $S$ into $S_0\cc\cxynv\cc S_1$ where $S_0\indep\bma$.
Let us also split $\bma$ into $T_0\cc\czwmv\cc T_1$ where $T_1\indep\cxynv$.
From \cref{can_move_intersection} we know that
$B\cc S \equiv \acb\cc\bma\cc S$.
We also know commands in $S_0$ and $\bma$ commute, and so this is equivalent to
$\acb\cc S_0\cc\bma\cc\cxynv\cc S_1$.
Expanding $\bma$ and swapping $T_1$ and $\cxynv$ we get
\[ \acb\cc S_0\cc T_0 \cc \czwmv \cc \cxynv\cc T_1\cc S_1. \]

@@

We know $\cxynv\nindep\czwmv$, and so from \cref{unrelated_is_independent}, $n\nunrel m$.
As the sequence works, $\czwmv\cc\cxynv$ can only be a construction or destruction pair.
As $\czwmv$ and $T_0$ originate from $B\setminus A$ and $B$ is simple, we also know there are
no commands on $m$ in $A\cap B$ or $T_0$,
and that there are no commands on $m$ in $S_0$ as $S_0\indep\bma$.

>>>>

\subsection{The Completeness of Reconciliation}

% Reconciliation is maximal
% -------------------------
% This is where we're using |D|=1

@@

\begin{myth}\label{rec_is_complete}
If $A$ and $B$ are simple sequences,
and we select a sequence $S\subset \amb$ so that
it contains a command $\cxynv$ for which $\cxynv\nindep\bma$,
then $B\cc S$ either breaks all filesystems $A$ and $B$ do not break
(that is, $S$ breaks all possible $\FS_B$ replicas),
or it changes a node that $B$ has already changed to a different value,
that is, it overrides a change in $\FS_B$.
\end{myth}
Such an override could occur if a given node was modified differently in
$A$ and $B$ (to different values but the same type), which is clearly
a conflict that should be marked for review
by a human or a different system.
\begin{proof}
The proof is similar to the ones we have seen above.
Without loss of generality, we assume that $\cxynv$ is the first command in $S$
that is not independent of $\bma$.
If so, we can split $S$ into $S_0\cc\cxynv\cc S_1$ where $S_0\indep\bma$.
Let us also split $\bma$ into $T_0\cc\czwmv\cc T_1$ where $T_1\indep\cxynv$.
From \cref{can_move_intersection} we know that
$B\cc S \equiv \acb\cc\bma\cc S$.
We also know that commands in $S_0$ and $\bma$ commute, and so this is equivalent to
$\acb\cc S_0\cc\bma\cc\cxynv\cc S_1$.
Expanding $\bma$ and swapping $T_1$ and $\cxynv$ we get
\[ \acb\cc S_0\cc T_0 \cc \czwmv \cc \cxynv\cc T_1\cc S_1. \]

@@

We know $\cxynv\nindep\czwmv$, and so from \cref{incomparable_is_independent}, $n\nunrel m$.
As the sequence works, $\czwmv\cc\cxynv$ can only be a construction or destruction pair.
As $\czwmv$ and $T_0$ originate from $B\setminus A$ and $B$ is simple, we also know there are
no commands on $m$ in $A\cap B$ or $T_0$,
and that there are no commands on $m$ in $S_0$ as $S_0\indep\bma$.
